Nail: A practical interface generator for data formats

### 2014-05-02 paper/osdi14/design.tex
Rewrote n_of. one paragraph over

\paragraph{Choices.}
If multiple structures can appear at a given position in a format, the programmer lists the options
along with a label for each in the \cc{choose} combinator. 
During parsing, Nail remembers the current input position and attempts each option in the order they
appear in the grammar. If an option fails, the parser backtracks to the initial position. If no
options succeed, the entire combinator fails. In the data model, choices are represented as tagged
unions.  The programmer has to be careful when options overlap, because if the programmer meant to
generate output for a choice, but the external representation is also valid for an earlier,
higher-priority option, the parser will interpret it as such. However, real data formats normally do
not have this overlap and we did not encounter it in the grammars we wrote.

@@

\paragraph{References.}
Rules allow for recursive grammars. To support recursive data types, we allow introduce the
reference combinator \cc{*}  that does not change the syntax of the external format described, but
introduces a layer of indirection, such as a reference or pointer, to the model data type.
The reference combinator does not need to used when another combinator, such as \cc{optional} or
\cc{many} already introduces indirection in the data type.
For example example,   arithmetic expressions consisting of sums and products can be parsed as follows:

\begin{verbatim}
expr = choose {
  PAREN = <uint8='('; *expr; uint8=')'>
  PRODUCT = sepBy1 uint8='*' expr
  SUM = sepBy1 uint8='+' expr
  INTEGER = many1 uint8 | '0' .. '9'
}
\end{verbatim} \cc{paren = < uint8="("; optional *paren; uint8=")" >}.

% To support recursive structures, we include a reference combinator \cc{*}, which
%   is syntactically equivalent to directly including the parser, but in the internal model adds a
%   level of indirection to the value. This allows recursive structures to be parsed, f

@@

Dependent fields are parsed just like other fields in the external format, but they are not stored
in the internal data type. Instead, the parser stores the value in a
temporary location, so that the field can be referenced by subsequent parsers.

When generating output, Nail visits a dependent field twice. First, while generating the other fields of a
structure, the generator reserves space for the dependent field in the output. Once
the dependent field goes out of scope, the generator  writes the
dependent field's value to this space.

Nail provides only one built-in combinator that uses dependent fields, \cc{n_of}, which is used
for count fields. For example, DNS labels are encoded as Pascal-style strings: a length followed by
the value of the label. The Nail grammar for DNS labels can represent
this pattern as:

{
\smaller[0.5]
\begin{verbatim}
  {
    @length uint8 | 1..64
    label n_of @length uint8
  }
\end{verbatim}
}

Other dependencies, such as offset fields or checksums, are not handled directly by combinators, but
through  transformations, as we describe next.

>>>>

\paragraph{Choices.}
If multiple structures can appear at a given position in a format, the programmer lists the options
along with a label for each in the \cc{choose} combinator. 
During parsing, Nail remembers the current input position and attempts each option in the order they
appear in the grammar. If an option fails, the parser backtracks to the initial position. If no
options succeed, the entire combinator fails. In the data model, choices are represented as tagged
unions.  The programmer has to be careful when options overlap, because if the programmer meant to
generate output for a choice, but the external representation is also valid for an earlier,
higher-priority option, the parser will interpret it as such. However, real data formats normally do
not have this overlap and we did not encounter it in the grammars we wrote.
An example is provided in Figure~\ref{fig:grammar-arith}.

@@

\paragraph{References.}
Rules allow for recursive grammars. To support recursive data types, we allow introduce the
reference combinator \cc{*}  that does not change the syntax of the external format described, but
introduces a layer of indirection, such as a reference or pointer, to the model data type.
The reference combinator does not need to used when another combinator, such as \cc{optional} or
\cc{many} already introduces indirection in the data type. An example is shown in Figure~\ref{fig:grammar-arith}.

\begin{figure}
\begin{verbatim}
expr = choose {
  PAREN = <uint8='('; *expr; uint8=')'>
  PRODUCT = sepBy1 uint8='*' expr
  SUM = sepBy1 uint8='+' expr
  INTEGER = many1 uint8 | '0' .. '9'
}
\end{verbatim} \cc{paren = < uint8="("; optional *paren; uint8=")" >}.
\label{fig:grammar-arith}
\caption{Grammar for sums and products of integers.}
\end{figure}
% To support recursive structures, we include a reference combinator \cc{*}, which
%   is syntactically equivalent to directly including the parser, but in the internal model adds a
%   level of indirection to the value. This allows recursive structures to be parsed, f

@@

Dependent fields are handled like other fields when parsing input, but their values are not stored
in the internal data type. Instead the value can be referenced by subsequent parsers and it
discarded when the field goes out of scope.
When generating output, Nail visits a dependent field twice. First, while generating the other fields of a
structure, the generator reserves space for the dependent field in the output. Once
the dependent field goes out of scope, the generator  writes the
dependent field's value to this space.

Nail provides only one built-in combinator that uses dependent fields, \cc{n_of}, which acts like
the \cc{many} combinator, except it represents an exact number, specified in the dependent field, of
repetitions, as opposed as many repetitions as possible. For example, DNS labels, which
are encoded as a length followed by
a value are described in line 38 of the DNS grammar.
Other dependencies, such as offset fields or checksums, are not handled directly by combinators, but
through  transformations, as we describe next.
